---
title: "Lab 5"
author: "Omima Hassan, Maureen McDonald"
date: "11/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
```


## *Processing the data*

1.
```{r}
diabetes = read.table('diabetes.txt', header=TRUE)
head(diabetes)
```

2.
```{r}
diabetes$frame[diabetes$frame == ''] = NA
diabetes$frame = droplevels(diabetes$frame)
```

3.
```{r}
diabetes_reduced = subset(diabetes, select = -c(id, bp.2s, bp.2d))
```


4.
```{r}
diabetes_clean = na.omit(diabetes_reduced)
```


## *Exploring and transforming data*
5.
```{r}
nrow(diabetes_clean)
ncol(diabetes_clean)
```

6.
* Using the log base along the x-axis. The downside to this is that the values are now smaller.  

7.
```{r}
diabetes_clean$glyhb_star = log(diabetes_clean$glyhb)
ggplot(diabetes_clean, aes(glyhb_star)) + geom_histogram() + labs(title = "Histogram of Glycosolated Hemoglobin", x = "Glycosolated Hemoglobin", y = "Amount in Patients")
```

8. -- Need to finish explanations of each summary statistic 
```{r}
diabetes_clean %>% group_by(gender) %>% summarise(mean.glyhb = mean(glyhb_star))

diabetes_clean %>% group_by(age) %>% summarise(mean.glyhb = mean(glyhb_star))

diabetes_clean %>% group_by(weight) %>% summarise(mean.glyhb = mean(glyhb_star))

diabetes_clean %>% group_by(height) %>% summarise(mean.glyhb = mean(glyhb_star))

```


* Gender and mean Glycosolated Hemoglobin. In terms of gender the mean amount of hemoglobin is about the same for males and females. 

* Age and mean Glycosolated Hemoglobin. In terms of age there isn't a clear relationship on whether mean amount of hemoglobin is higher for those who are older although some values could indicate that.

* Weight and mean Glycosolated Hemoglobin. In terms of weight those who weigh more don't seem to have a higher amount of hemoglobin than those who way less, again outliers do exist where values for mean hemoglobin for those who weigh less are more than those who weigh more.

* Height and Glycosolated Hemoglobin. In terms of height those who are taller have a higher mean value for glycosolated hemoglobin.

## *Visualizations*
10.
```{r}
diabetes_clean%>% 
  group_by(frame,location) %>% 
  summarize(mean.glyhb_star = mean(glyhb_star,na.rm=TRUE)) %>% 
  ggplot(aes(frame, mean.glyhb_star, color=location))+ 
  geom_point()
```

11.
```{r}
ggplot(diabetes_clean,aes(glyhb_star, ratio))+ geom_point(size=.5)+ geom_smooth() + labs(title = "Glycosolated Hemoglobin vs Ratio", x = "Glycosolated Hemoglobin", y = "Ratio")
```

```{r}
ggplot(diabetes_clean,aes(glyhb_star, bp.1s))+ geom_point(size=.5)+ geom_smooth() + labs(title = "Glycosolated Hemoglobin vs First Systolic Blood Pressure", x = "Glycosolated Hemoglobin", y = "First Systolic Blood Pressure")
```


```{r}
ggplot(diabetes_clean,aes(glyhb_star, age))+ geom_point(size=.5)+ geom_smooth() + labs(title = "Glycosolated Hemoglobin vs Age", x = "Glycosolated Hemoglobin", y = "Age")
```

```{r}
ggplot(diabetes_clean,aes(gender,glyhb_star))+ geom_boxplot() + labs(title = "Glycosolated Hemoglobin vs Gender", x = "Gender", y = "Glycosolated Hemoglobin")
```

```{r}
ggplot(diabetes_clean,aes(glyhb_star, hip))+ geom_point(size=.5)+ geom_smooth() + labs(title = "Glycosolated Hemoglobin vs Hip (inches)", x = "Glycosolated Hemoglobin", y = "Hip (inches)")
```

```{r}
ggplot(diabetes_clean,aes(glyhb_star, weight))+ geom_point(size=.5)+ geom_smooth() + labs(title = "Glycosolated Hemoglobin vs weight (pounds)", x = "Glycosolated Hemoglobin", y = "Weight (pounds)")
```

12.
```{r}
ggplot(diabetes_clean,aes(y=hip,x=waist,alpha = 0.5)) + geom_point(shape=1, size=.4)+ facet_wrap(~frame) + labs(title='Waist vs. Hip by Frame',x='Waist (in.)', y='Hip (in.)')
```

## *Messy data*

13.
* Gather brings variables that could be collapsed together, to the same column.
* Spread takes columns that have several forms of information, like date and time, that can be divided into different rows.

14.
* Gather and spread are not exact complements of each other because you cannot constantly go back and forth from spread and gather. As you gather the information it must all become one. If you spread the information it forgets about the original one, creating a completely new column.


## *Regression models*
15.
```{r}
fit = lm(glyhb_star ~stab.glu + age + waist + ratio+ factor(frame),data=diabetes_clean)
summary(fit)
```

* Based on these results, we can conclude that these variables are not significant because of how small the P-value is. We would then choose to reject the null hypothesis based on that. To improve this model, more variables could be added. 
Based on question 12, we think a linear model would be the right approach. If you look back at that question, you can see that a positive linear relationship exists.


16.
* For an increase of stabilized glucose, the fit increases by 0.0035.
* For an increase of age by 1 year, the fit increases by 0.0034.
* For an increse of waist by 1 inch, the fit increases by 0.0048.
* For an increase of ratio, the fit increases by 0.0219.
* For an increase of factor(frame)medium, the fit increases by 0.0309.
* For an increase of factor(frame)small, the fit increases by 0.3750.


17.
* The true value is the average  amount of glycosolated hemoglobin when a persons frame = large and stabilized glucose is 0 and age, waist, and ratio are not all equal to 0 but one of them is.

* $\hat{Y}$ =  0.8330897 + 90*stab.glu + 35*age + 30*waist + 5.1*ratio + 0.0131840*FrameSmall  


18.
* Inference is when you are given a data set and have to get an output from the date.
* Prediction is getting another data set and seeing how the old data set helps to build a model.

19.
* Advantages- can see if positive or negatively relationship, parametric shape, easy to interpret and fit
* Disadvantages- parametric shape, only will show something if linear, assumes the values are normally distributed

## *Reflection*
20.
* We thought that data science is the study of analyzing information and knowledge from sets of data. We both agree that this is true. The amount of statistics that is used in data science was surprising to us both. The most challenging thing is when R doesn't want to work. We spent many nights in the basement of Pearson yelling at our screens because R kept crashing. Visualizations are our favorite thing to do. You can do a lot of cool things with it.
